Induction

### 2023-10-29 induction/induction.tex
strong induction

  If $n = k$, then by definition $f(k) = 2f(k-1) - f(k-2)$.\footnote{Note that $k-2$ and $k-1$ are natural numbers because $k > 1$ in this case. We eliminated $k = 0$ and $k = 1$ in the previous step for this step to work.}
  Since $k - 2 < k$ and $k - 1 < k$, the induction hypothesis says that $f(k-2) = 2(k-2) + 1 = 2k - 3$ and that $f(k-1) = 2(k-1) + 1 = 2k - 1$.
  Thus,
  \begin{align}
    f(k) &= 2f(k-1) - f(k-2)\\
         &= 2(2k - 1) - (2k - 3)\\
         &= 4k - 2 - 2k + 3\\
         &= 2k + 1
  \end{align}
\end{proof}

@@

>>>>

  If $n = k$, then by definition $f(k) = 2f(k-1) - f(k-2)$.
  Since $k - 2 < k$ and $k - 1 < k$, the induction hypothesis says that $f(k-2) = 2(k-2) + 1 = 2k - 3$ and that $f(k-1) = 2(k-1) + 1 = 2k - 1$.
  Thus,
  \begin{align}
    f(k) &= 2f(k-1) - f(k-2)\\
         &= 2(2k - 1) - (2k - 3)\\
         &= 4k - 2 - 2k + 3\\
         &= 2k + 1
  \end{align}
\end{proof}

@@

\begin{prop}\label{prop:lincomb}
  For any $n \in \dN$, if $n \geq 2$ then $n$ is a linear combination of $2$ and $3$, i.e., there are natural numbers $i$ and $j$ so that $n = 2i + 3j$.
\end{prop}

Again, we need to strengthen the induction hypothesis.

\begin{lem}\label{lem:lincomb-strong}
  For any $m \in \dN$ and $n \in \dN$, if $n < m$ then if additionally $n \geq 2$ then $n$ is a linear combination of $2$ and $3$.
\end{lem}
\begin{proof}
  $P(m)$ is $\forall n.\,(n < m \imp (n \geq 2 \imp \exists i.\,\exists j.\,n = 2i + 3j))$.

  (Base case): $P(0)$ is $\forall n.\,(n < 0 \imp (n \geq 2 \imp \exists i.\,\exists j.\,n = 2i + 3j))$.
  Let $n \in \dN$.
  Assume $n < 0$.
  This is a contradiction.

  (Induction step): Let $k \in \dN$.
  Assume $P(k)$, i.e.,
  \[
    \forall n.\,(n < k \imp (n \geq 2 \imp \exists i.\,\exists j.\,n = 2i + 3j))
  \]
  We need to prove $P(k+1)$, i.e.,
  \[
    \forall n.\,(n < k + 1 \imp (n \geq 2 \imp \exists i.\,\exists j.\,n = 2i + 3j))
  \]
  Let $n \in \dN$.
  Assume $n < k + 1$ and $n \geq 2$, so $k \geq 2$.
  If $k = 2$ then $n$ has to be $2$, which can be expressed as $2 \cdot 1 + 3 \cdot 0$.
  If $k = 3$ then $n$ has to be $2$ or $3$.
  We know how to express $2$, and $3$ can be expressed as $2 \cdot 0 + 3 \cdot 1$.
  For $k \geq 4$, if $n < k$ then the induction hypothesis gives us what we want.
  If $n = k$, then consider $k - 2$.
  The induction hypothesis tells us that there are natural numbers $a$ and $b$ so that $k - 2 = 2a + 3b$.
  We can then express $n$ as $n = k = k - 2 + 2 = 2(a + 1) + 3b$.
\end{proof}

\begin{proof}[Proof of \cref{prop:lincomb}]
  Let $n \in \dN$.
  Assume $n \geq 2$.
  The result follows immediately from \cref{lem:lincomb-strong}.
\end{proof}
